---
title: "Mapping Segregation in R"
output: html_document
---
R **segregation** package allows us to calculate segregation measures.

- The Mutual Information Index (M)

- Theil’s Information Index (H)

- The index of Dissimilarity (D)

More information about **segregation** package can be found in the [vignette.](https://elbersb.github.io/segregation/articles/segregation.html)
and [r documentation](https://cran.r-project.org/web/packages/segregation/segregation.pdf)
Other practices can be found [here.](https://elbersb.github.io/segregation/)


### Required packages

```{r, results="hide", warning=FALSE, message=FALSE}
library(segregation)  # to calculate segregation index
library(tidycensus)   # to obtain demographic data from the Census
library(tidyverse)    # to clean the data
library(viridis)      # to map with colors
library(tigris)       # to obtain TIGER/Line shapefiles from the Census
#If you have not installed the Census API Key, you need to do this.
#census_api_key("YOUR KEY GOES HERE")
```

We first try with Illinois. We could expect that the Chicago Metropolitan area would be most segregated. Get the demographic data of Illinois using `get_acs` in **tidycensus** package.

```{r}
il_data <- get_acs(geography = "county", 
                   variables = c(white = "B03002_003",
                                 black = "B03002_004",
                                 asian = "B03002_006",
                                 hispanic = "B03002_012"),
                   state = "IL")
```

### Calculate segregation index
(a) `dissimilarity` calculates the total segregation between group and unit.

```{r}
il_data %>% 
     filter(variable %in% c("black", "white")) %>% 
     dissimilarity(group = "variable", unit = "GEOID", weight = "estimate")
```

(b) `mutual_total` calculates total segregation for M and H.
```{r}
il_data %>% 
     mutual_total(group = "variable",
                  unit = "GEOID",
                  weight = "estimate")
```

(c) `mutual_local` calculates local segregation, decomposed by units or groups. Here our `unit` is GEOID (geographical unit) and our `group` is race variable. We will use this segregation index for mapping. 

```{r}
il_local_seg <- il_data %>% 
     mutual_local(group = "variable",
                  unit = "GEOID",
                  weight = "estimate", 
                  wide = TRUE)
```

Then we make the Illinois local segregation data we had above into a merged data calling shapefiles from the Census using `counties` in **tigris** package. We merge it by GEOID. 

```{r, results="hide", warning=FALSE, message=FALSE}
il_local_seg_geom <- counties("IL", cb=TRUE)%>%
     left_join(il_local_seg, by = "GEOID")
```

### Plotting segregation data
Finally plot the local segregation index at the county level using `ggplot`. 

```{r}
ggplot(il_local_seg_geom, aes(fill=ls)) +
     geom_sf(color = NA) +
     coord_sf(crs = 3435) +
     scale_fill_viridis_c() +
     theme_void() +
     labs(fill = "Local/segregation index IL")
```

The map shows the level of segregation at county level in Illinois, but unlike our expectation that Chicago area would be the most segregated, the map displays the opposite. Maybe we need to consider smaller geographical units, such as tracts or cities, because some of these units tend to have more diverse racial groups.  

Let's try with Illinois but at the tract level following almost similar steps. 

```{r, results="hide", warning=FALSE, message=FALSE}
# tract level

il_tract_data <- get_acs(geography = "tract", 
                         variables = c(white = "B03002_003",
                                       black = "B03002_004",
                                       asian = "B03002_006",
                                       hispanic = "B03002_012"),
                         state = "IL")
il_local_seg_tract <- il_tract_data %>% 
     mutual_local(group = "variable",
                  unit = "GEOID",
                  weight = "estimate", 
                  wide = TRUE)
il_local_seg_tract_geom <- tracts("IL", cb=TRUE)%>%
     left_join(il_local_seg_tract, by = "GEOID")
ggplot(il_local_seg_tract_geom, aes(fill=ls)) +
     geom_sf(color = NA) +
     coord_sf(crs = 3435) +
     scale_fill_viridis_c() +
     theme_void() +
     labs(fill = "Local/segregation index IL at the tract level")
```

We now have a better map, higher (brighter area) segregation in Chicago area. 


This time let's focus only on Cook County whose county seat is Chicago.


```{r, results="hide", warning=FALSE, message=FALSE}
# county level : Cook county, IL

cook_data <- get_acs(geography = "tract",
                     variables = c(white = "B03002_003",
                                   black = "B03002_004",
                                   asian = "B03002_006",
                                   hispanic = "B03002_012"),
                         state = "IL",
                         county = "Cook County")

cook_local_seg <- cook_data %>% 
     mutual_local(group = "variable",
                  unit = "GEOID",
                  weight = "estimate", 
                  wide = TRUE)

cook_local_seg_geom <- tracts(state = "IL", county = "Cook County", cb=TRUE)%>%
     left_join(cook_local_seg, by = "GEOID")

ggplot(cook_local_seg_geom, aes(fill=ls)) +
     geom_sf(color = NA) +
     coord_sf(crs = 3435) +
     scale_fill_viridis_c() +
     theme_void() +
     labs(fill = "Local/segregation index of Cook County at the tract level")
```

Let's do one more example with Harris County of Texas whose largest city is Houston.


```{r, results="hide", warning=FALSE, message=FALSE}
# county level : Harris county, TX

harris_data <- get_acs(geography = "tract", 
                       variables = c(white = "B03002_003",
                                     black = "B03002_004",
                                     asian = "B03002_006",
                                     hispanic = "B03002_012"),
                       state = "TX",
                       county = "Harris County")

harris_local_seg <- harris_data %>%
                                mutual_local(group = "variable",
                                             unit = "GEOID",
                                             weight = "estimate", 
                                             wide = TRUE)

harris_local_seg_geom <- tracts(state = "TX", county = "Harris County", cb=TRUE)%>%
     left_join(harris_local_seg, by = "GEOID")

ggplot(harris_local_seg_geom, aes(fill=ls)) +
     geom_sf(color = NA) +
     coord_sf(crs = 3435) +
     scale_fill_viridis_c() +
     theme_void() +
     labs(fill = "Local/segregation index of Harris County at the tract level")
```






